---
title: "IntroBasics-JH"
author: "Justin Hollis"
date: "2023-01-21"
output: html_document
editor_options: 
  chunk_output_type: console
---
Load the Penguins Dataset
```{r}

data("penguins")
```


1.  Why don't the following lines of code work? Tweak each one so the code runs

    ```{r}
    myList[["Fruit"]]
    ```

    ```{r}
    penguins$flipper_length_mm
    ```

    ```{r}
    penguins%>%
      filter(island =='Dream')
    ```


2.  How many species are in the `penguins` dataset? What islands were the data collected for? (Note: the `unique()` function might help)

```{r}
length(unique(penguins$species))
unique(penguins$island)

```


3.  Use indexing to create a new data frame that has only 3 columns: species, island and flipper length columns, and subset all rows for just the 'Dream' island.

```{r}
subset <- penguins%>%  
  select(species, island, flipper_length_mm)%>%
  filter(island == "Dream")

```


4.  Use indexing and the `mean()` function to find the average flipper length for the *Adelie* species on Dream island. (Note: explore the `mean()` function and how to deal with NA values).
```{r}
mean(subset$flipper_length_mm[subset$species == "Adelie"])
```


6.3.2 Exercises

1.  Reorder the variables in `penguins` so that `year` is the first column followed by the rest (Hint: look into the use of `everything()`).
```{r}
yearfirst <- penguins %>% 
  relocate(year, .before = species)


```


2.  Create a new column called 'size_group' where individuals with body mass greater than the overall average are called 'large' and those smaller are called 'small'.
```{r}

avg <- mean(penguins$body_mass_g, na.rm = TRUE)
 yearfirst %>% 
  mutate(size_group = case_when(avg < body_mass_g ~ "Large",
                              avg > body_mass_g ~ "Small"))
 
  
```


3.  Find out which year for each species were individuals on average the largest according to body mass. ????? 

```{r}
penguins %>% 
group_by(year, species) %>% 
  summarise(mean(penguins$body_mass_g, na.rm = TRUE), year)
```


4.  You want to filter data for years that are *not* in a vector of given years, but this code doesn't work. Tweak it so that it does. (Yes, you could just filter year to equal 2007 in this case but there is a trouble-shooting lessons here).
    ```{r}
    penguins %>% 
      filter(year !%in% c(2008, 2009))
    ```

### 7.1 Exercises

1.  Make a barplot showing the average flipper length for each species.
```{r calculating the mean flipper length for each species}
meanflip <- penguins %>% 
group_by(species) %>% 
summarise(mean = mean(flipper_length_mm, na.rm = TRUE))
```


```{r graphing the mean flipper length for each species}
ggplot(data = meanflip)+
  geom_col(mapping = aes(x = species, y = mean , fill = species), width = 0.5)
```

2.  Make a scatter plot of bill length compared to bill depth but only for observations on the Dream island.
```{r grouping by observation on dream island}
 dream <- penguins%>%
      filter(island =='Dream')

```
```{r plotting bill length vs bill depth }

ggplot(data = dream)+
  geom_point(mapping = aes(x = bill_length_mm, y = bill_depth_mm, color = species))

```




