---
output: html_document
editor_options: 
  chunk_output_type: console
---


```{r}
source("lessons/setup.R")
```


#Import Colorado Counties with Tigris 
```{r}
counties <- counties(state="CO")
```


# Import Roads for Larimer County 
```{r}
roads <- roads(state = "CO", county = "Larimer")
```

```{r}
tmap_mode("view")
qtm(counties)+
  qtm(roads)
```



#look at the class of counties 
```{r}
class(counties)
```

#point data 
```{r}
poudre_points <- data.frame(name = c("Mishawaka", "Rustic", "Blue Lake Trailhead"),
                            long = c(-105.35634, -105.58159, -105.85563),
                            lat = c(40.68752, 40.69687, 40.57960))


```


#Covert to spatial

```{r}
poudre_points_sf <- st_as_sf(poudre_points, coords = c("long", "lat"), crs = 4326)
qtm(poudre_points_sf)
```


#rasterdata
```{r}

elevation <- get_elev_raster(counties, z=7)

qtm(elevation)
tm_shape(elevation)+
  tm_raster(style ="cont" , title = "Elevation (m)")
```



#the terra package 

```{r}
elevation <-rast(elevation)

names(elevation) <- "Elevation"

```


#Check Projections 

```{r}
st_crs(counties)
crs(counties) == crs(elevation)
```

#project elevation layer 
```{r}
elevation_prj<-- terra::project(elevation, counties)
```


#crop elevation to counties extent 
```{r}
elevation_crop <- crop(elevation, ext(counties))
qtm(elevation_crop)
```



#Read and Write Spatial Data 
```{r}
write_sf(counties, "data/counties.shp")
```



#Save raster data 
```{r}
writeRaster(elevation_crop, "data/elevation.tif")
```



#save .Rdata
```{r}
save(counties, roads, file= "data/spatial_objects.Rdata")
rm(counties, roads)

```


#1.  Filter out the `counties` data set to only include Larimer, Denver, and Pueblo counties.

```{r}
countsub <- counties %>%  
 filter(NAME == "Denver" | NAME == "Larimer" | NAME == "Pueblo")

```


#2.  Make a map of the `counties` data colored by county area. Make a second map of counties colored by their total area of water.

```{r}
tmap_mode("view")
qtm(countsub)
```


#3.  Make a barplot comparing the elevation of your 3 points in the Poudre Canyon (note: explore the `extract()` function in the `terra` package).
cannot get the points to plot 

```{r pull elevation points}
elevation_points <- get_elev_point(poudre_points_sf)
```
```{r barplot}
ggplot(data = elevation_points)+
  geom_col(mapping = aes(x = name, y = elevation , fill = name), width = 0.5)+
  labs(x = "", y = "Elevation (m)", fill = "")
  
```


#4.  Why are there 4 features in our Poudre Canyon Highway variable instead of 1?


